---
title: "P3-Dimgray-Assignment"
author: "Jared Teo, Nicholas Poh, Oliver Choy, WuJie, Nicole, Kurt"
format: html
---

# Required Packages

```{r}
#| label: required-packages
#| message: false

library(tidyverse)
library(gt)
library(palmerpenguins)
library(ggbeeswarm)
library(ggmosaic)
library(ggplot2)
library(dplyr)
library(readr)

```

# Data
```{r}
#| label: inflation-rate
#| message: false

inflation_rate <- read_csv("M212911.csv", skip = 10)

inflation_rate <- inflation_rate[1:152,]

inflation_rate 
```

```{r}
#| label: inflation-rate-plot
#| fig-height: 6
#| fig-width: 10
#| message: false
# Filter the data
filtered_data <- inflation_rate |>
  filter(`Data Series` %in% 
  c("Food", "Food Excl Food Serving Services", "Clothing & Footwear","Housing & Utilities","Household Durables & Services","Health Care","Transport","Communication", "Recreation & Culture", "Education", "Miscellaneous Goods & Services", "All Items")) |>
  select(`Data Series`, `2023 Jan`, `2023 Feb`,`2023 Mar`, `2023 Apr`,`2023 May`, `2023 Jun`, `2023 Jul`, `2023 Aug`,`2023 Sep`, `2023 Oct`,`2023 Nov`, `2023 Dec`)

filtered_month_data <- filtered_data |>
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filtered_month_data <- filtered_month_data |>
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filtered_month_data, aes(month, inflation_rate, group = `Data Series`)) +
  geom_point(aes(color = month, size = inflation_rate)) +
  geom_path(aes(color = month))+
  geom_text(
    aes(label = `Data Series`),
    slice_max(filtered_month_data, month), 
    vjust = 1,
    nudge_y = 2
  )

```

# Filter and plot "All Items" data
```{r}
#| label: filter-all-items
#| fig-height: 4
#| 
#| message: false
#| 
filer_all_items <- filtered_data %>%
  filter(`Data Series` == "All Items") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filer_all_items <- filer_all_items %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filer_all_items, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'All Items' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```


# Filter and plot "Food" data
```{r}
#| label: filter-Food
#| fig-height: 4
#| message: false
#| 
filter_food <- filtered_data %>%
  filter(`Data Series` == "Food") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_food <- filter_food %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_food, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Food' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Food Excl Food Serving Services" data
```{r}
#| label: filter-Food-ex
#| fig-height: 4
#| message: false
#| 
filter_food_ex <- filtered_data %>%
  filter(`Data Series` == "Food Excl Food Serving Services") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_food_ex <- filter_food_ex %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_food_ex, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Food Excl Food Serving Services' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Clothing & Footwear" data
```{r}
#| label: filter-cloths
#| fig-height: 4
#| message: false
#| 
filer_cloths <- filtered_data %>%
  filter(`Data Series` == "Clothing & Footwear") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filer_cloths <- filer_cloths %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filer_cloths, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Clothing & Footwear' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Housing & Utilities" data
```{r}
#| label: filter-house
#| message: false
#| 
filter_house <- filtered_data %>%
  filter(`Data Series` == "Housing & Utilities") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_house <- filter_house %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_house, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Housing & Utilities' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Household Durables & Services" data
```{r}
#| label: filter-household
#| message: false
#| 
filter_household <- filtered_data %>%
  filter(`Data Series` == "Household Durables & Services") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_household <- filter_household %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_household, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Household Durables & Services' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Healthcare" data
```{r}
#| label: filter-healthcare
#| message: false
#| 
filter_healthcare <- filtered_data %>%
  filter(`Data Series` == "Health Care") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_healthcare <- filter_healthcare %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_healthcare, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Healthcare' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Transport" data
```{r}
#| label: filter-transport
#| message: false
#| 
filter_transport <- filtered_data %>%
  filter(`Data Series` == "Transport") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_transport <- filter_transport %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_transport, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Transport' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Communication" data
```{r}
#| label: filter-communication
#| message: false
#| 
filter_communication <- filtered_data %>%
  filter(`Data Series` == "Communication") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_communication <- filter_communication %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_communication, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Communication' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Recreation & Culture" data
```{r}
#| label: filter-recreation
#| message: false
#| 
filter_recreation <- filtered_data %>%
  filter(`Data Series` == "Recreation & Culture") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_recreation <- filter_recreation %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_recreation, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Recreation & Culture' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Education" data
```{r}
#| label: filter-education
#| message: false
#| 
filter_education <- filtered_data %>%
  filter(`Data Series` == "Education") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_education <- filter_education %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_education, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Education' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

# Filter and plot "Miscellaneous Goods & Services" data
```{r}
#| label: filter-miscellaneous
#| message: false
#| 
filter_miscellaneous <- filtered_data %>%
  filter(`Data Series` == "Miscellaneous Goods & Services") %>%
  pivot_longer(cols = -`Data Series`, names_to = "month", values_to = "inflation_rate")

# Convert month to proper date format
filter_miscellaneous <- filter_miscellaneous %>%
  mutate(month = as.POSIXct(strptime(paste(month, "01"), "%Y %b %d")))

ggplot(filter_miscellaneous, aes(x = month, y = inflation_rate, group=1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Line graph of 'Miscellaneous Goods & Services' inflation percentage change through 2023",
    x="Month", 
    y="Value",
    caption = "Source: Department of Statistics Singapore") +
  theme_classic()
```

